---
title: "ECO395M Data Mining: HW 3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
dengue = read.csv("C:/Users/musab/OneDrive/Documents/ECON MA/Data mining/HW3/dengue.csv")
library(rpart)
library(randomForest)
library(gbm)
library(rpart.plot)
library(Metrics)
library(rsample)
library(pdp)
library(dplyr)
```

```


# Question 1
Jacob will take care of this.

# Question 2

```{r, message=FALSE, echo=FALSE}
# Exclude rows with missing values
dengue_clean <- na.exclude(dengue)

# Convert 'city' and 'season' to factors
dengue_clean$city <- as.factor(dengue_clean$city)
dengue_clean$season <- as.factor(dengue_clean$season)

set.seed(123) # Seed for reproducibility
split <- initial_split(dengue_clean, prop = 0.9)
train <- training(split)
test <- testing(split)

```

```{r, message=FALSE, echo=FALSE}
# Fit the initial CART model
cart_fit <- rpart(total_cases ~ ., data = train, 
                  control = rpart.control(cp = 0.002, minsplit = 20))

# Plot the un-pruned CART model
rpart.plot(cart_fit, digits = 3, type = 4, extra = 1)

# Prune the CART model
prune_func <- function(tree) {
  fit_details <- as.data.frame(tree$cptable)
  optimal_cp <- with(fit_details, max(CP[xerror - xstd <= min(xerror)]))
  prune(tree, cp = optimal_cp)
}
cart_pruned <- prune_func(cart_fit)

# Calculate RMSE for the pruned CART model
pred_cart <- predict(cart_pruned, test)
rmse_cart <- sqrt(mean((pred_cart - test$total_cases)^2))
```
```{r, message=FALSE, echo=FALSE}
# Fit the Random Forest model
rf_fit <- randomForest(total_cases ~ ., data = train, ntree = 200)

# Calculate RMSE for the Random Forest model
pred_rf <- predict(rf_fit, test)
rmse_rf <- sqrt(mean((pred_rf - test$total_cases)^2))
```

```{r, message=FALSE, echo=FALSE}
# Fit the GBM model
gbm_fit <- gbm(total_cases ~ ., data = train, 
               distribution = "gaussian", n.trees = 300, 
               interaction.depth = 3, shrinkage = 0.1, 
               cv.folds = 5, n.cores = 1)

# Determine the optimal number of trees
opt_trees <- gbm.perf(gbm_fit, method = "cv")

# Calculate RMSE for the GBM model
pred_gbm <- predict(gbm_fit, test, n.trees = opt_trees)
rmse_gbm <- sqrt(mean((pred_gbm - test$total_cases)^2))
```

#The GBM model's performance plot reveals how it learns over time, with the training error steadily decreasing, reflecting a good fit on training data. The validation error's dip and subsequent plateau suggest an optimal number of trees at the dashed line, beyond which the model may start to overfit.

```{r, message=FALSE, echo=FALSE}
cat("RMSE for Pruned CART Model: ", rmse_cart, "\n")
cat("RMSE for Random Forest Model: ", rmse_rf, "\n")
cat("RMSE for GBM Model: ", rmse_gbm, "\n")
```
#The Random Forest model outperforms its counterparts with an RMSE of 23.67, making it the most accurate for predicting dengue cases in this dataset, while the CART and GBM models show higher errors, at 28.40 and 25.93 respectively.

```{r, message=FALSE, echo=FALSE}
if (rmse_rf == min(c(rmse_cart, rmse_rf, rmse_gbm))) {
  # PDP for specific_humidity
  pdp_humidity <- partial(rf_fit, pred.var = "specific_humidity", train = train, grid.resolution = 10)
  plot(pdp_humidity, main = "Partial Dependence Plot for Specific Humidity", xlab = "Specific Humidity", ylab = "Predicted Dengue Cases")
  
  # PDP for precipitation_amt
  pdp_precip <- partial(rf_fit, pred.var = "precipitation_amt", train = train, grid.resolution = 10)
  plot(pdp_precip, main = "Partial Dependence Plot for Precipitation Amount", xlab = "Precipitation Amount", ylab = "Predicted Dengue Cases")
  
  # PDP for tdtr_k
  pdp_tdtr <- partial(rf_fit, pred.var = "tdtr_k", train = train, grid.resolution = 10)
  plot(pdp_tdtr, main = "Partial Dependence Plot for Diurnal Temperature Range", xlab = "Temperature Range", ylab = "Predicted Dengue Cases")
} else {
  cat("The Random Forest model did not perform best. Adjust the code to plot PDPs for the best-performing model.")
}
```

#The Partial Dependence Plots suggest that higher specific humidity may increase dengue cases, while diurnal temperature range shows a possible decrease in cases with greater fluctuations. Precipitation amount, however, doesn't show a clear impact, indicating that other factors might be at play or that its relationship to dengue cases isn't straightforward. These insights can help focus preventive measures on more impactful environmental factors.

# Question 3
Daniel will take care of this.
# Question 4
I have started working on this one already.
## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
