---
title: "ECO395M Data Mining: HW 3"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
```{r, include=FALSE}
dengue = read.csv("C:/Users/musab/OneDrive/Documents/ECON MA/Data mining/HW3/dengue.csv")
library(rpart)
library(randomForest)
library(gbm)
library(rpart.plot)
library(Metrics)
library(rsample)
library(pdp)
library(dplyr)
```

```


# Question 1
Jacob will take care of this.
# Question 2
Musab will take care of this.
# Question 3
Daniel will take care of this.
# Question 4
I have started working on this one already.
## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
